---
title: "Presentation Ninja"
subtitle: "⚔<br/>with xaringan"
author: "Yihui Xie"
institute: "RStudio, Inc."
date: "2016/12/12 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---


# Data types

There are several data types in R. The four main types of data.

 - Lists
 - Vectors
 - Matrices
 - Data frames

```{r data types}
###Vectors contain similar elements. It can be numeric or characters#####
###The simplest way to create a vector is to use the function c () combine####
b = c(1, 3, 2, 6)
class(b)
```

```{r}
### Alternatively, we can create a vector of the characters with the name of European cities####
cities = c('Madrid', 'London', 'Ostrava', 'Pekin', 'Barcelona')
class(cities)
```

---
# Data types (cont..)


```{r}
###We can also check the length of the vector by using the lenght () function
length(b)  # It contains 4 observation, 4 elements. It determines the number of elements in the vector

### In the case of a character vector, we need to use the nchar () function to count the number of characters of each element
nchar(cities)  #
```
---

## Data frames

--
- Sometimes data require more complex storage than simple vectors and thankfully R provides a host of data structures. The most common are *the data.frame*, matrix, list, and vectors. Perhaps, the data frame will be most familar to anyone who has used Excel spreadsheet.

- The structure of the data.frame is similar to an Excel spreadsheet. Each column represents a variable, and each row is an observation. Data frames might contain different types of data (numeric, character, logical, etc)


```{r}
#Load the sample dataset: USArrests
data(USArrests)
head( USArrests,5)
# You can also make quick changes in the dataset with the fix () command
fix(USArrests) # editing your dataset. Similar to the edit command in Stata
```

---
## Data frames (cont..)

### Basic operations with data.frames

```{r}
nrow(USArrests) ## Number of rows. Number of observations = 50 US states
ncol(USArrests) ## Number of variables in the dataset.  4 variables
## Alternatively, 
dim(USArrests) ## Number of rows and columns of our dataset. with dimension function. 
```
---
## Data frames (cont..)
### Basic operations with data.frames

```{r}
names(USArrests) ## names of the variables in the dataset
colnames(USArrests)  ## checking the names of the variables

#These functions return a character vector listing the columns 
names(USArrests)[1] ##We pick up the first element of the vector

# The same logic applies to the rows of our dataset

rownames(USArrests) ## Checking the row names
rownames(USArrests)[50]  ## Selecting Wyoming state
```
---
Thanks