---
title: "Lab 5: Basic plots and ggplot package."
output: github_document
---

### Basic plots using R base

- We will produce basic graphs in R using the dataset (auto.mpg). I am providing the slides and the full code which shows how to build up different plots with the basic syntax in R. It is assumed that you are familiar with descriptive statistics.


### Reading the dataset

```{r}
#Check the structure of the dataset
data.auto = read.csv('../Labs/data/auto-mpg.csv')
head(data.auto)
attach(data.auto)
```

### Number of cylinders as a categorical variable (factor in R)

```{r}
str(cylinders)
cylinders = factor(cylinders,
levels = c(3,4,5, 6, 8),
labels = c('3cyl', '4cyl', '5cyl', '6cyl', '8cyl'))
head(cylinders)
```


### Histogram

```{r}
hist(acceleration, probability = TRUE)
```

Let us try to refine the histogram by changing some options

```{r}
hist(acceleration, 
     main = 'Histogram of accelaration',
     xlab = 'Acceleration',
     col = 'blue')
```

We can also choose the number of break points in hist() as follows:


```{r}
hist(acceleration, 
     main = 'Histogram of accelaration',
     xlab = 'Acceleration',
     col = 'blue',
     breaks = 16)
```


### Fitting a density function

```{r}
hist(mpg, breaks = 16, prob = T)
lines(density(mpg))
```




###  Box plot
```{r}
boxplot(mpg, xlab = "Consumption")
boxplot(mpg ~ model_year, xlab = "Millas por Galeón (por año)")
boxplot(mpg ~ cylinders, xlab = "Consumo por Número de cilindros")
boxplot(data.auto)
```

### Simple Scatter Plot

```{r}
plot(mpg ~ acceleration, col ='green')
```

### Fitting a simple regression model

```{r}
plot(mpg ~ horsepower)
linearmodel <- lm(mpg ~ horsepower)
abline(linearmodel)
```

### Scatterplot matrices

We can also create scatter plot matrices with the pairs() function.

```{r}
pairs(~mpg + horsepower + weight + acceleration, col ='green')
```

## Combination of plots with par ()

```{r warning=FALSE}
old.par <- par()
old.par
par(mfrow = c(1,2))
with(data.auto, {
  plot(mpg ~ weight, main = "Weight vs Consumption")
  plot(mpg ~ acceleration, main = "Acceleration vs Consumption")
}
)
par(old.par)
```





## Introduction to ggplot

R provides excellent graphing capabilities with pacakges such as **lattice** and **ggplot2**.

For more details on using R Markdown see <http://rmarkdown.rstudio.com>. We will briefly describe some simple graphs using base graphics from the previous unit and then show their counterparts in ggplot2.


## What is ggplot?

An R package designed to create plots based on a theory of the grammar of the graphics

## Why ggplot instead of base R?

* nice defaults

* easy faceting

* more natural syntax

* can switch chart types more easily

## Reading and checking the structure of the dataset

```{r}
data.auto = read.csv('../Labs/data/auto-mpg.csv')
head(data.auto)
attach(data.auto)
```

## ggplot syntax

The basic structure for ggplot2 starts up with the **ggplot** function. It can take additional arguments or fewer but we will stick with that for the time being.

In addition to this, we add layers (options) with the plus sign. We will discuss geometric layers with points, or lines with the following functions

1. geom_point ()
2. geom_line ()
3. geom_histogram ()


##Histograms and densities with ggplot2

We will plot the distribution of acceleration using ggplot2.

```{r warning = FALSE}
library(ggplot2)
ggplot(data = data.auto) + geom_histogram(aes(x= acceleration))
```


## Categorical variables

```{r}
ggplot(data =data.auto) + geom_bar(aes(x= cylinders))
table(cylinders)
```

##Scatterplots

```{r}
g = ggplot(data.auto, aes(x=mpg, y =weight)) + geom_point()
g + ggtitle('Scatter plot') + xlab('Miles/US Galon') + ylab('Weight of the vehicle per 1000lbs')
```


## Boxplots
```{r}
ggplot(data.auto, aes(x= factor(cylinders), y= mpg)) + geom_boxplot()
```

## Violin plots
```{r}
ggplot(data.auto, aes(x=factor(cylinders), y =weight)) + geom_point() + geom_violin()
```
