---
title: "Lab 6: Logistic Regression. Economic Analysis with R"
output: github_document
---
## Introduction

* Logistic regression or logit model is employed to model dichotomous outcome variable (dependent variable). 
* It is also a very popular classification algorithm in machine learning.  

 * At the end of this unit, students will be able to:

    *  know what does exactly logistic regression
    *  become familiar with the syntax of R logistic regression
     * apply the logit model in different contexts
     

# Applications of logistic regression 

* Health care analysis
* Geographic image processing
* To make predictions about something we employ logistic regression



## Basic syntax

* It is quite similar to regression models using the lm () command. We employ the *glm ()* function that has the following basic arguments

\begin{equation}
glm(formula, data, family)
\end{equation}

* The formula indicates the relationship between variables. The dataset is the dataframe containing the information of the variables, and the family is the R object to specify the details of the model. In the case of logistic regression, its value is binomial.


# Performance of the logistic regression

* We need to test the performance of our model by using different metrics.
   * Akaike criteria
   * Null deviance
   * Confusion matrix


# Titanic dataset

We will be using the Titanic dataset from Kaggle. Let us try to download it, and summarize the dataset


```{r}
fitting.data.raw = read.csv("../Labs/data/train.csv", header = TRUE, na.strings = c(''))
str(fitting.data.raw)
summary(fitting.data.raw)
```


## Missing values treatment

```{r}
install.packages('mice', repos='http://cran.us.r-project.org')
library(mice)
```

## Pattern of missing values
```{r}
md.pattern(fitting.data.raw)
```


## Visualizing missing values

```{r}
install.packages('VIM',repos='http://cran.us.r-project.org')
library(VIM)
```

##Graphical analysis

```{r}
aggr_titanic = aggr(fitting.data.raw, col = mdc(1:2), numbers = TRUE, sortVars = TRUE)
```


## Subset, drop variables with lots of missing values, it can be imputed

```{r}
fitting.data.raw.copy1 = subset(fitting.data.raw, select = c(2, 3,5, 6, 7, 8, 10, 12))
str(fitting.data.raw.copy1)
```
Alternativaley, with the column number or column names.

```{r}
fitting.data.raw.copy2 = fitting.data.raw[,c(2,3,5,6,7,8, 10, 12)]
str(fitting.data.raw.copy2)
```

## Imputed missing values for Age with the mean

```{r}
install.packages('Hmisc',repos='http://cran.us.r-project.org')
library(Hmisc)
fitting.data.raw.copy1$Age = impute(fitting.data.raw.copy1$Age, mean)
summary(fitting.data.raw.copy1)

```
## Categorical variables ( factors in R: sex, and embarked)

```{r}
sapply(fitting.data.raw.copy1, class)
```

## Remove NAs for one variable (Embarked)
```{r}
fitting.data.raw.copy1 = fitting.data.raw.copy1[!is.na(fitting.data.raw.copy1$Embarked),]
summary(fitting.data.raw.copy1)
```


# Practical example

```{r}
# load the dataset and create a small dataframe
temperatura <-c(66,70,69,68,67,72,73,70,57,63,70,78,67,53,67,75,70,81,76,79,75,76,58)
defecto <-c( 0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 0, 0, 1, 0, 0, 0, 0, 0, 0, 0, 0, 1)
aux <-matrix(c(temperatura,defecto),ncol = 2)
colnames(aux) <- c('temperatura','defecto')
datos<-data.frame(aux)
```

```{r}
# Question. Count the number of correct/wrong

table(datos$defecto)
round(prop.table(table(datos$defecto)), 2)
```

```{r}
# Graphical representation
colores <- NULL
colores[datos$defecto == 0] <- "green"
colores[datos$defecto == 1] <- "red"
plot(datos$temperatura, datos$defecto, pch = 21, bg = colores, xlab = "Temperatura", ylab = "Prob. defecto")
legend("bottomleft", c("No defecto", "Si defecto"), pch = 21, col = c("green", "red"))
```


```{r}
# Estimate a logistic regression model where the dependent variable is defecto and the explanatory variable is the temperature
reg <- glm(defecto ~ temperatura, data = datos, family = binomial)
summary(reg)
```


# References

*R for everyone: Advanced Analytics and Graphics. Chapter 20. Generalized linear models. Logistic regression. Addison Wesley Data & Analytics Series.