---
title: "Lab 5: Using the ggplot package."
output: github_document
---


## Introduction

R provides excellent graphing capabilities with pacakges such as **lattice** and **ggplot2**.

For more details on using R Markdown see <http://rmarkdown.rstudio.com>. We will briefly describe some simple graphs using base graphics from the previous unit and then show their counterparts in ggplot2.


## What is ggplot?

An R package designed to create plots based on a theory of the grammar of the graphics

## Why ggplot instead of base R?

* nice defaults

* easy faceting

* more natural syntax

* can switch chart types more easily

## Reading and checking the structure of the dataset

```{r}
data.auto = read.csv('../Labs/data/auto-mpg.csv')
head(data.auto)
attach(data.auto)
```

## ggplot syntax

The basic structure for ggplot2 starts up with the **ggplot** function. It can take additional arguments or fewer but we will stick with that for the time being.

In addition to this, we add layers (options) with the plus sign. We will discuss geometric layers with points, or lines with the following functions

1. geom_point ()
2. geom_line ()
3. geom_histogram ()


##Histograms and densities with ggplot2

We will plot the distribution of acceleration using ggplot2.

```{r warning = FALSE}
library(ggplot2)
ggplot(data = data.auto) + geom_histogram(aes(x= acceleration))
```


## Categorical variables

```{r}
ggplot(data =data.auto) + geom_bar(aes(x= cylinders))
table(cylinders)
```

##Scatterplots

```{r}
g = ggplot(data.auto, aes(x=mpg, y =weight)) + geom_point()
g + ggtitle('Scatter plot') + xlab('Miles/US Galon') + ylab('Weight of the vehicle per 1000lbs')
```


## Boxplots
```{r}
ggplot(data.auto, aes(x= factor(cylinders), y= mpg)) + geom_boxplot()
```

## Violin plots
```{r}
ggplot(data.auto, aes(x=factor(cylinders), y =weight)) + geom_point() + geom_violin()
```
