---
title: "Lab 1: First Steps in R. Economic Analysis with R"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## First steps
1. Open RStudio on yours or a group member's laptop
2. Check RStudio version (version command, mine )
2. Click File > New File > R script. 

You will see a blank section of screen in the top-left of your RStudio window. This is where you will write your first R script. This is similar to write a do file in Stata. Notice that R code is saved as a `.r file`.

### Four basic areas on your screen

1. The is the `script` editor as in Stata
2. The bottom left is the `console` (you can use R interactively by typing commands as in Stata)
3. The upper right is where you see the list of objects in your workspace. History is also there (you know at aytime what you did type in your console)
4. The bottom right is where you find the available packages, graphical output, your file structure, and help.  

### Help

Getting help with R is easy. To get further information about the syntax of any R function,  just type a `?` before the function name. See the example below

```{r}
# regression function (fitting linear models by OLS)
?lm
```


If you don't know exactly the name of the fuction, just use double question marks

```{r}
??cluster
help.search("cluster")
apropos("cluster")
```


Finally, the structure of the help for each R function is:

1. Description
2. Usage
3. Arguments
4. Value
5. References

### Console
The bottom-left of the screen has a tab called "Console". This is basically a very fancy calculator.

Try the basic calculator by typing something like
```{r}
2+2
```

Or even something fancier like
```{r}
sqrt(pi)
b <- 1.4156
sqrt(b)
```


### Objects and basic R

R is a oriented object language. For R, everything is an object, and each object has its own name.

1. To remove any particular object from the current session employ `rm()`
2. To list all objects in your workspace use `ls()`
3. To remove all objects from your workingspace just type `rm(list =ls())`

```{r}
x<-2 # Creating an object
ls()
rm(list = ls())
```


Lastly, most of the them when using variables in a dataframe, you refer to them by using the `$`symbol. This can be avoided by using the `attach` command.

Just as an example, let us load one of the built in datasets in R, and compute the gas consumption per mile.

```{r}
data(mtcars)
mean(mtcars$mpg) # Notice that we get access to the variables with the dollar symbol

attach(mtcars)
mean(mpg) # there is no need to use the dollar symbol
```


### Packages
R makes extensive use of third-party packages. We will not get into the details right now, but for this class, you will need to install a few of these. Installing packages is quite easy. Type the following two lines of code at the very top of your script:

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
install.packages("tidyverse", repos='http://cran.us.r-project.org')
install.packages("skimr", repos='http://cran.us.r-project.org')
install.packages("wooldridge", repos='http://cran.us.r-project.org')
```

You've just installed two basic packages. We can also check it out by using the following syntax

```{r}
installed.packages()
```




Basically, you have downloaded them onto your laptop. Just like with other software on your computer, you only need to do the installation once. However, you still need to tell R that you will be using the packages (by using the `library`command). Add the following two lines of code to your script (below the first two lines you wrote). Notice how there are no quotation marks inside the parenthese this time.

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(tidyverse) # one of the most famous R pacakge for data manipulation
library(skimr)  # for data exploration
library(wooldridge) # it refers to Wooldrige's undergraduate texbook
```

Whenever you start over a new Rstudio session, you need to load the approrpiate libraries.

In RStudio, you can also install packages by using the R interface. You will find it at Tools -> Install Package, and there you will get a pop--up window to type the package you want to install


### Running a script
To execute the script, click on the word "Source" in the top-right corner of the top-left window pane. This will take what is in your script and automatically send it to the console (as if you typed it directly into the console)

To save the script, click on the disk icon at the top of your script pane (but not the disk icon at the very top of RStudio). Name your script `ICL1_XYZ.R` where `XYZ` are your initials.

### Commenting
Now, put a hashtag (#) in front of the first two lines of code in your script, like so:
```{r}
#install.packages("tidyverse")
#install.packages("skimr")
#install.packages("wooldridge")
```

The hashtag is how you tell R not to run the code in your script. This is known as "commenting" your code.

At the very top of your script, type the names of your group members with a hashtag in front.

### Citation R and packages 

If you need to cite R, just use

```{r}
citation()
```

if for instance, you need to cite one of the packages, then use

```{r}
citation(package = "ggplot2")
```

