---
title: "Tutorial 2: Descriptive statistics"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
library(shiny)
```

## Measures of central tendency

Generate your own dataset with the list of numbers and compute the mean, median, and the trimmed mean (removing 10 % of the observations)

```{r dataset, exercise = TRUE}
datos <- c(-15, 2, 3, 4, 5, 6, 7, 8, 9, 12)
print(datos)
mean(datos)
median(datos)
mean(x = datos, trim = 0.1) # Notice that we remove 10 %
```

Try to compute the mean using a simple code for the previous numeric vector

```{r mean, exercise = TRUE}
# We divide the output of summation by the number of elements 
```

```{r mean-solution}
datos <- c(-15, 2, 3, 4, 5, 6, 7, 8, 9, 12)
sum(datos)/10
```


Let us calculate the mode, the most frequent value, what we can do is to count the number of times on which each gender appears on the sample. The core R packages do not have a function for calculating the mode.

```{r mode, exercise = TRUE, message=FALSE}
library(funModeling)
head (heart_disease, 3)
# Compute the most frquent gender in the sample
```

```{r mode-solution}
table(heart_disease$gender) # Males are more frequent in our sample
```


## Measures of variability

We want to know how spread out are the data. Let us compute the range, interquartile range, and variance.

```{r variability, exercise = TRUE}
data(mtcars)
min(mtcars$mpg)
max(mtcars$mpg)
range(mtcars$mpg)
```


```{r quantile, exercise = TRUE}
# Let use the quantile function to compute the median, and check it out with the median function
```

```{r quantile-solution}
quantile (mtcars$mpg, probs = 0.5)
median(mtcars$mpg)
```

```{r quantile2, exercise = TRUE}
#let us compute the quantiles (1, 2, and 3 quartile)

```

```{r quantile2-solution}
quantile (mtcars$mpg, probs = c(0.25, 0.50, 0.75))
```


```{r variance, exercise = TRUE}
# Let us compute the variance and standard deviation

```

```{r variance-solution}
var(mtcars$mpg)
sqrt(var(mtcars$mpg))
```

### Skew and kurtosis


*Perhaps more helpfully, it might be useful to point out that the psych
package contains a skew() function that you can use to calculate skewness.*

```{r skew, exercise = TRUE}
# Let us compute the skewness and kurtosis
```


```{r skew-solution}
install.packages("psych", repos = "https://cran.cnr.berkeley.edu/")
library(psych)
skew(mtcars$mpg)
kurtosi(mtcars$mpg)
```

### Quiz

*You can include any number of single or multiple choice questions as a quiz. Use the `question` function to define a question and the `quiz` function for grouping multiple questions together.*

Some questions to verify that you understand the purposes of various base and recommended R packages:

```{r quiz}
quiz(
  question("Which package contains functions for installing other R packages?",
    answer("base"),
    answer("tools"),
    answer("utils", correct = TRUE),
    answer("codetools")
  ),
  question("Which of the R packages listed below are used to create plots?",
    answer("lattice", correct = TRUE),
    answer("tools"),
    answer("stats"),
    answer("grid", correct = TRUE)
  )
)
```

